---
title: Model estimates in multiple cruises.
date: "Compiled at `r format(Sys.time(), '%Y-%m-%d %H:%M:%S', tz = 'UTC')` UTC"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r global_options, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(fig.width=14, fig.height=8, echo=TRUE, eval=TRUE, cache=TRUE,
                      warning=FALSE, message=FALSE,
                      cache.lazy = FALSE)

## Load packages
library(tidyverse)
library(knitr)
library(here)
library(ggplot2)
library(flowmix)


## Setup some plotting details
coul <- colorRampPalette(RColorBrewer::brewer.pal(8, "RdYlBu"))(25)
source("02-helpers.R")

## Some extra stuff, from: https://github.com/haozhu233/kableExtra/issues/265
options(kableExtra.auto_format = F)
```

 
```{r directories}
## Setup locations
base = "02-estimate"
here::i_am("02-estimate.Rmd")
knitr::opts_chunk$set(fig.path = here::here("data", base, 'figures/'))
outputdir = here::here("data", base)
datadir = here::here("data", base)
outputdir_cytograms = here::here("data", "01-cytograms")
if(!dir.exists(outputdir)) dir.create(outputdir)
```

# Summary

This document compiles and showcases the estimated models -- results from
large-scale computation that occurred in a server.

The experiments were done using these settings:

* Number of clusters: default to 10.
* Size of time blocks: blocks of 5 hours
  + Block 1 is: 1,...,5 
  + Block 2 is 6,..., 10. And so forth.
  + Then, CV is done by using blocks {1, 6, 11, ..} as fold 1, {2, 7, 12, ..} as
    fold 2, and so on.
* The "radius" for the cluster means over time: a value of 1 seems to work well.

Here are some notes on computation:

* For KM1508, one iteration costs 2 or 2.5 seconds per iteration.
* On 500 cores, entire CV can be done in 15-30 minutes.
* Subsampling will take (100 subsamples x 30 min = 50 hrs = 1 day).

# Experiments

## Run scripts on server 

For now, the scripts are placed on the USC CARC server. Two of the most relevant
files:

* `~/scripts/flowmixapp/run-main.slurm` which calls
* `~/scripts/flowmixapp/main.R`.

Note to self: these scripts will eventually be made available directly in this
repository.

## Download model results

ONLY download the `summary.RDS` file.

```{sh, eval = FALSE}
all_cruise_id="KM1508 KM1512 KM1513 KOK1515 KM1518 KM1601 KM1602 KOK1604 KOK1606 KOK1607 KOK1608 KOK1609 MGL1704 KM1708 KM1709 KM1713 KM1717  KOK1801 KM1802 KOK1803 FK180310-1 FK180310-2"
for cruise_id in 
do
  echo $cruise_id
  from=sangwonh@discovery.usc.edu:projdir/output/flowmixapp/data/02-estimate/$cruise_id/summaries/*.RDS
  to=/home/sangwonh/repos/flowmixapp/data/02-estimate/$cruise_id/summaries/.
  mkdir -p  $to
  rsync -auv $from $to
done
```

# Cluster labels

First, here are some hand-labeled cluster labels:

```{r cluster-labels, echo = FALSE}
## All mappings 
maps =  list(KM1508  = c(pro = c(10,NA), small = 9,  syn = 7,  pico1_ = c(6  ), pico2_ = c(2,NA), pico3_ = c(NA), pico4_ = c(NA), croco =NA, bead = 1, rest1_ = c(4,NA), rest2_ = c(8), bg = c(3,5)),
             KM1512  = c(pro = c(10,NA), small = 9,  syn =NA,  pico1_ = c(7  ), pico2_ = c(8,NA), pico3_ = c(NA), pico4_ = c( 1), croco = 3, bead = 6, rest1_ = c(2,NA), rest2_ = c(4), bg = c(5)  ),
             KM1513  = c(pro = c(8, NA), small =10,  syn =NA,  pico1_ = c(7  ), pico2_ = c(6,NA), pico3_ = c(NA), pico4_ = c(NA), croco = 5, bead = 1, rest1_ = c(3,NA), rest2_ = c(9,4), bg = c(2)),
             KOK1515 = c(pro = c( 7, 3), small = 9,  syn =NA,  pico1_ = c(10 ), pico2_ = c(2,NA), pico3_ = c(NA), pico4_ = c(NA), croco = 5, bead =NA, rest1_ = c(1,NA), rest2_ = c(8), bg = c(6,4)),

             KM1518  = c(pro = c(10,NA), small = 6,  syn = 3,  pico1_ = c(7  ), pico2_ = c(1,NA), pico3_ = c(NA), pico4_ = c(NA), croco =NA, bead = 4, rest1_ = c(2,NA), rest2_ = c(9), bg = c(8,5)),
             KM1601  = c(pro = c( 4,10), small = 7,  syn = 5,  pico1_ = c(3  ), pico2_ = c(2,NA), pico3_ = c(NA), pico4_ = c(NA), croco =NA, bead = 6, rest1_ = c(1,NA), rest2_ = c(9), bg = c(8)),
             KM1602  = c(pro = c( 7, 9), small = 8,  syn =NA,  pico1_ = c(4  ), pico2_ = c(3,NA), pico3_ = c(NA), pico4_ = c(NA), croco =NA, bead = 6, rest1_ = c(2, 1), rest2_ =c(10), bg = c(5)),
             KOK1604  = c(pro = c(10,NA), small = 2,  syn =NA, pico1_ = c(6  ), pico2_ = c(1,NA), pico3_ = c( 9), pico4_ = c(NA), croco =NA, bead = 4, rest1_ = c(7, 5), rest2_ = c(8), bg = c(3)),

             KOK1606 = c(pro = c( 9,NA), small =NA,  syn =10,  pico1_ = c(5  ), pico2_ = c(8,NA), pico3_ = c(7),  pico4_ = c(NA), croco =NA, bead =  4, rest1_ = c(1,NA), rest2_ = c(6), bg = c(3,2)),
             KOK1607 = c(pro = c(10,NA), small = 8,  syn = 9,  pico1_ = c(4  ), pico2_ = c(3,NA), pico3_ = c(NA), pico4_ = c(NA), croco =NA, bead =  5, rest1_ = c(6, 7), rest2_ = c(1), bg = c(2)),
             KOK1608 = c(pro = c(10,NA), small =NA,  syn =NA,  pico1_ = c(3  ), pico2_ = c(5, 1), pico3_ = c(NA), pico4_ = c(NA), croco = 2, bead =  4, rest1_ = c(6, 7), rest2_ = c(9), bg = c(8)),
             KOK1609 = c(pro = c( 7,NA), small =NA,  syn = 5,  pico1_ = c(3  ), pico2_ = c(1,NA), pico3_ = c(NA), pico4_ = c(NA), croco = 4, bead = NA, rest1_ = c(9, 2), rest2_ = c(6), bg = c(10,8)),

             ## Non-pilot cruises go here.
             MGL1704 = c(pro = c(10,NA), small =NA,  syn = 3,  pico1_ = c(9  ), pico2_ = c(1,NA), pico3_ = c( 6), pico4_ = c(NA), croco =NA, bead = NA, rest1_ = c(4, 7), rest2_ = c(8), bg = c(5, 2)),
             KM1708  = c(pro = c( 8,NA), small =NA,  syn =NA,  pico1_ = c(9  ), pico2_ = c(7,NA), pico3_ = c(NA), pico4_ = c( 5), croco = 3, bead =  2, rest1_ = c(6,NA), rest2_ = c(10),bg = c(4, 1)),

             KM1709  = c(pro = c( 5,NA), small =NA,  syn =NA,  pico1_ = c(6  ), pico2_ = c(3, 7), pico3_ = c(NA), pico4_ = c(NA), croco = 4, bead = NA, rest1_ = c(9, 8), rest2_ = c(2), bg = c(10,1)),
             KM1713  = c(pro = c( 9,NA), small =NA,  syn =NA,  pico1_ = c(8 ), pico2_ = c(4, 7), pico3_ = c(NA), pico4_ = c(NA), croco = 1, bead = NA, rest1_ = c(2, 3,  5), rest2_ = c(10), bg = c(6)), 

             KM1717  = c(pro = c( 8,NA), small =NA,  syn = 9,  pico1_ = c(6  ), pico2_ = c(7,NA), pico3_ = c( 1), pico4_ = c(NA), croco = 2, bead = NA, rest1_ = c(3, 5), rest2_ = c(4), bg = c(NA,10)),
             KM1802 = c(pro = c( 9,NA), small =10,  syn = 4,   pico1_ = c(5  ), pico2_ = c(6,NA), pico3_ = c(NA), pico4_ = c(NA), croco =NA, bead =  2, rest1_ = c(1,NA), rest2_ = c(8), bg = c(3, 7)),
             KOK1801 = c(pro = c( 9,NA), small =NA,  syn =NA,  pico1_ = c(4  ), pico2_ = c(3, 6), pico3_ = c(NA), pico4_ = c(NA), croco =NA, bead =  2, rest1_ = c(1, 5), rest2_ = c(10),bg = c(7, 8)),

             KOK1803 = c(pro = c( 10,8), small = 7,  syn = 5,  pico1_ = c(3  ), pico2_ = c(1,NA), pico3_ = c(NA), pico4_ = c(NA), croco =NA, bead =  4, rest1_ = c(2,NA), rest2_ = c(9),bg = c(6,NA)), ## not run yet.

        "FK180310-1" = c(pro = c( 8,NA), small =NA,  syn = 9,  pico1_ = c(10 ), pico2_ = c(7, 2), pico3_ = c(NA), pico4_ = c(NA), croco =NA, bead =  5, rest1_ = c(6, 4), rest2_ = c(3 ),bg = c(1,NA)),
        "FK180310-2" = c(pro = c( 9,NA), small =NA,  syn = 5,  pico1_ = c(10 ), pico2_ = c(1,NA), pico3_ = c(NA), pico4_ = c(NA), croco =NA, bead =  3, rest1_ = c(2, 7), rest2_ = c(8 ),bg = c(4, 6))
             )
             
mapmat = maps %>% bind_rows(.id = "cruise_id") %>% as_tibble()
mapmat_long = mapmat %>% pivot_longer(!cruise_id)
numclust = 10
mapmat %>% select(-cruise_id) %>% as.matrix() %>% apply(1, function(a){
  assertthat::assert_that(all(1:numclust == sort(a)))
}) 
knitr::kable(mapmat) 
```


# Estimated $\pi$ and $\mu$ {.tabset}

Next, we'll make videos of model, and plots \& tables of coefficients.

```{r, load-data, results = 'asis', fig.width=15, fig.height=5}
all_cruise_id = c(c("KM1508", "KM1512", "KM1513",
                    "KOK1515", "KM1518"),
                  c("KM1601", "KM1602", "KOK1604", "KOK1606"),
                  c("KOK1607", "KOK1608", "KOK1609"), 
                  c("MGL1704", "KM1708", "KM1709", "KM1713", "KM1717", "KOK1801", "KM1802"), 
                  c("KOK1803",  "FK180310-1", "FK180310-2"))

## Too little data in 1805
## c("KM1712"), ## Need to download covariates, bin properly and run again.

numclust = 10

## Loop over all cruise IDs
for(cruise_id in all_cruise_id){
  cat('## Model', cruise_id,' \n')

  ## Cluster labels
  if(cruise_id %in% mapmat$cruise_id){
    one_mapmat_long = mapmat_long %>% filter(cruise_id == !!(cruise_id)) %>% arrange(value) %>% na.omit()
    stopifnot((one_mapmat_long %>% pull(value)) == (1:numclust))
    cluster_labels = one_mapmat_long %>% pull(name)
  } else {
    cluster_labels = NULL
  }

  ## Load model
  bestres = load_model(cruise_id)

  ## Load cytogram data.
  filename = paste0(cruise_id, "-datobj.RDS")
  ybin_obj = readRDS(file = file.path(outputdir_cytograms, cruise_id, "data", filename))
  ylist = ybin_obj$ylist
  countslist = ybin_obj$biomass_list
  
  ## Create individual frames
  dir.create(file.path(file.path(outputdir, cruise_id, "viz")))
  TT = length(ylist)
  for(tt in 1:TT){
    g = plot_2d_threepanels(bestres, ylist, countslist, tt = tt,
                            labels = cluster_labels,
                            colours = c("white", "blue"),
                            cruise_id = cruise_id)

    ## Save one for the books
    if(tt == 1){
      ## oneplotfile = file.path(datadir, cruise_id, "viz", paste0(tt, ".jpeg"))
      ## ggsave(file = oneplotfile, width = 12, height = 4, units = "in", dpi = 300, g)
      plot(g)
    }

    ## Save one for the books
    filename = paste0("labeled-", tt, ".png")
    plotfile = file.path(file.path(outputdir, cruise_id, "viz"), filename)
    if(file.exists(plotfile)) next
    ggsave(file = plotfile,
           width = 12, height = 4,
           units = "in", dpi = 300, g)
  }

  ## Make video
  videofile = file.path(outputdir, cruise_id, "viz", "3d-model.mp4")
  if(!file.exists(videofile)){
    cmd_change_directory <- paste0("cd ", file.path(file.path(outputdir, cruise_id, "viz")))
    cmd_make_movie <- "ffmpeg -y -framerate 3 -i labeled-%d.png  3d-model.mp4"
    system(paste0(cmd_change_directory, "; ", cmd_make_movie))
  }

  ## Plot cluster probabilities
  times = names(ylist)
  plot_prob_ggplot(bestres, cluster_labels, times) %>% print()
  
  ## Embed video (not doing this anymore, due to large file size)
  if(FALSE){
    cat(paste0('<video width="1280" height="720" controls>  <source src="',
               videofile,
               '" type="video/mp4"> </video>'))
  }

  ## ## Instead, we'll just embed one frame
  ## cat(paste0("![](", file.path(".", 
  ##                              "data",
  ##                              "02-estimate",
  ##                              cruise_id,
  ##                              "viz",
  ##                              "1.jpeg"),
             ## ")"))
  cat('\n\n')
}
```
 

# Estimated $\alpha and $\beta$ coefficients {.tabset}

```{r viz_coef, fig.width=12, fig.height=12, results='asis'}
## These are the original /ordered/ labels
orig_cluster_labels = mapmat %>% colnames()

## Setup (assuming |ylist| exists) 
cytogram_dimnames = ylist[[1]] %>% colnames()

for(cruise_id in all_cruise_id){
  cat('## Tables', cruise_id,' \n')

  ## Cluster labels
  one_mapmat_long = mapmat_long %>% filter(cruise_id == !!(cruise_id)) %>% arrange(value) %>% na.omit()
  if(nrow(one_mapmat_long) == 0) next
  stopifnot((one_mapmat_long %>% pull(value)) == (1:numclust))
  cluster_labels = one_mapmat_long %>% pull(name)

  ## Load model
  bestres = load_model(cruise_id)##, cluster_labels)
  
  ## Plot and print alpha coefficients
  p = print_alpha(bestres, type = "plot", cluster_labels, orig_cluster_labels)
  print(p)
  tab = print_alpha(bestres, type = "table", cluster_labels, orig_cluster_labels) ##%>% print()
  print(tab)
 
  ## Plot and print beta coefficients
  p = plot_beta(bestres, cluster_labels, orig_cluster_labels, cytogram_dimnames)
  plot(p)
  p = print_beta(bestres, cluster_labels, orig_cluster_labels, cytogram_dimnames)
  print(p)

  cat('\n\n')
}
```


# Some diagnostics {.tabset}

You can never be too careful! Two things are shown, for each cruise:

* The matrix of CV scores (shown as a heatmap).

* Sparsity of all models (shown as tables).
 
  * The row names show the values of $\lambda_\alpha$.
  * The column names show the values of $\lambda_\alpha$.

```{r diagnostics, fig.width=8, fig.height=8, results='asis'}
for(cruise_id in all_cruise_id){
  cat('##', cruise_id,' \n')

  ## Show the CV
  summarydir = file.path(outputdir, cruise_id, "summaries")
  cvres = readRDS(file.path(summarydir, "summary-orig.RDS"))
  mat = cvres$cvscore.mat
  mat %>% drawmat_precise() %>% plot() ##col = c("blue", "blue", "red"))

  ## Print the sparsity of the model coefficients.
  alphadf = betadf = cvres$cvscore.mat
  for(ii in 1:10){
    for(jj in 1:10){
      obj = cvres$bestreslist[[paste0(ii, "-", jj)]]
      betas = do.call(cbind, lapply(obj$beta, function(beta)beta[-1,]))
      alpha =obj$alpha[ ,-1]
      betadf[ii,jj] = length(betas) - sum(betas==0)
      alphadf[ii,jj] = length(alpha) - sum(alpha==0)
    }
  }
  knitr::kable(alphadf, caption = "Number of nonzero alpha coeffs") %>% print() ## title "alpha"
  knitr::kable(betadf, caption = "Number of nonzero beta coeffs")  %>% print()## title "beta"
  cat('\n\n')
}
```


# Subsampled models

Download the results from the server:

```{sh, eval=FALSE}
all_cruise_id="KM1508 KM1512 KM1513 KOK1515 KM1518 KM1601 KM1602 KOK1604 KOK1606 KOK1607 KOK1608 KOK1609 MGL1704 KM1708  KM1712 KM1713 KM1717  KOK1801 KM1802 KOK1803 FK180310-1 FK180310-2"
for cruise_id in all_cruise_id
do
	echo $cruise_id
	from=sangwonh@discovery.usc.edu:projdir/output/flowmixapp/data/02-estimate/$cruise_id/summaries/*
	to=/home/sangwonh/repos/flowmixapp/data/02-estimate/$cruise_id/summaries/.
	mkdir -p  $to
	rsync -auv $from $to
done
```

Now, we'll load the summaries and aggregate the information from all experiments
-- original *and* subsampling replicates. 

This creates summary files named e.g.  `KM1601-grand-summary.RDS`.

```{r grand-summarize, eval = FALSE}
all_cruise_id = c("KM1508",
                  "KM1512",
                  ## "KM1513", ## This has fold 4 missing altogether; why is this?
                  "KOK1515",
                  "KM1518",
                  "KM1601",
                  "KM1602",
                  "KOK1604",
                  "KOK1606",
                  "KOK1607",
                  "KOK1608",
                  "KOK1609", ##"KOK1608a", , "KOK1609a"
                  ## c("KM1712"), ## Need to download covariates, bin properly and run again.
                  "MGL1704", "KM1708",
                  "KM1709", ## No subsampling yet
                  "KM1713", ## No subsampling yet
                  "KM1717", "KOK1801", "KM1802", ## These are new.
                  "KOK1803",  "FK180310-1", "FK180310-2") ## "KM1805" ## Too little data in 1805

for(cruise_id in all_cruise_id){
  tryCatch({
  subsample_summarize(cruise_id = cruise_id,
                      save = TRUE,
                      nsim = 100,
                      outputdir = "~/projdir/output/flowmixapp/data/02-estimate",
                      outputdir_cytograms = "~/projdir/output/flowmixapp/data/01-cytograms",
                      verbose = TRUE)
  }, error = function(e){
    print("error occurred!")
    return(NULL)
  })
}
```

## Beta coefficients and stability {.tabset}

```{r beta-stab, results="asis"}
varnames = c("diam", "chl", "pe")
numclust = 10
dimdat = 3
for(cruise_id in all_cruise_id){

  ## Read in the summary over all subsampled models.
  filename = paste0(cruise_id, "-grand-summary.RDS")
  summarydir = file.path(outputdir, cruise_id, "summaries")
  if(!file.exists(file.path(summarydir, filename))) next
  obj = readRDS(file = file.path(summarydir, filename))

  ## Cluster labels
  one_mapmat_long = mapmat_long %>% filter(cruise_id == !!(cruise_id)) %>% arrange(value) %>% na.omit()
  stopifnot((one_mapmat_long %>% pull(value)) == (1:10))
  cluster_labels = one_mapmat_long %>% pull(name)

  for(iclust in 1:numclust){
    ## cat("###", cruise_id, paste0("clust-", iclust))
    cat("###", cruise_id, cluster_labels[iclust])

    threetables = lapply(1:dimdat, function(idim){
      pipeline = function(a, varname){
        a %>% as.data.frame() %>%  magrittr::set_colnames(varname) %>% 
        tibble::rownames_to_column(var="variable")  %>% as_tibble()
      }
      stab = obj$beta_freq %>% .[iclust, ,idim] %>% pipeline("stab")
      coef = obj$origres$beta %>% .[[iclust]] %>% .[,idim] %>% pipeline("est")
      combined = full_join(stab, coef) %>%
        arrange(-stab) %>%
        mutate(stab = print_sparsify(stab),
               est = print_sparsify(est)) %>%
        filter(variable != "intp") %>%
        magrittr::set_colnames(c("variable", paste0(varnames[idim], "_stab"),
                                 paste0(varnames[idim], "_est")))
      return(combined)
    })

    ## Combine and kable-ify
    threetables %>% do.call(cbind, .) %>% knitr::kable(caption="beta coefficients") %>% print()
      cat(fill = TRUE)
  }
}
```


## Alpha coefficients and stability {.tabset}

```{r alpha-stab, results = "asis"}
numclust = 10
for(cruise_id in all_cruise_id){

  ## Read in the summary over all subsampled models.
  filename = paste0(cruise_id, "-grand-summary.RDS")
  summarydir = file.path(outputdir, cruise_id, "summaries")
  if(!file.exists(file.path(summarydir, filename))) next
  obj = readRDS(file = file.path(summarydir, filename))

  ## Cluster labels
  one_mapmat_long = mapmat_long %>% filter(cruise_id == !!(cruise_id)) %>% arrange(value) %>% na.omit()
  stopifnot((one_mapmat_long %>% pull(value)) == (1:10))
  cluster_labels = one_mapmat_long %>% pull(name)

  for(iclust in 1:numclust){
    ## cat("###", cruise_id, paste0("clust-", iclust))
    cat("###", cruise_id, cluster_labels[iclust])
    
    ## Obtain and combine coefficient estimates and stability estimates
    pipeline = function(a, varname){
      a %>% as.data.frame() %>%  magrittr::set_colnames(varname) %>% 
        tibble::rownames_to_column(var="variable")  %>% as_tibble()
    }
    stab = obj$alpha_freq[,iclust] %>% pipeline("stab")
    coef = obj$orig$alpha[iclust,] %>% pipeline("est")
    combined = full_join(stab, coef) %>% arrange(-stab)
    
    ## Sparsify
    combined = combined %>% mutate(stab = print_sparsify(stab),
                                   est = print_sparsify(est)) %>%
      filter(variable != "intp") 
    
    ## Kable-ify
    combined %>% kable(caption="alpha coefficients") %>% print()
    cat(fill = TRUE)
  } 
}
```


# Make three data frames

Now, we'll be making several "long format" data tables -- of the data and of the
model estimates -- to use for further data analysis.

The objects are called:

* `coef_long`: $\alpha$ and $\beta$ coefficient values, as well as stability estimates.

* `pred_long`: $\mu$ and $\pi$ estimates.

* `sigma_long`: Covariance ($\Sigma$) estimates.

* `X_long`: Environmental covariates.

* `dat_long`: Binned cytometry data.


**Coefficients (and stability).** The columns are:

  + cruise
  + "variable": covariate (can also be "intercept")
  + population
  + type ("beta_diam", "beta_chl", "beta_pe", "alpha")
  + coefficient (the value)
  + stability (the value)



```{r gather-subsampling}
objlist = sapply(all_cruise_id, function(cruise_id){

  ## Read in the summary over all subsampled models.
  filename = paste0(cruise_id, "-grand-summary.RDS")
  summarydir = file.path(outputdir, cruise_id, "summaries")
  if(!file.exists(file.path(summarydir, filename))){
    print(cruise_id)
    print("didn't have grand summary file")
    return(NULL)
  }
  obj = readRDS(file = file.path(summarydir, filename))
  return(obj)

}, simplify = FALSE, USE.NAMES = TRUE)
```


```{r coef_long}
coef_long = lapply(all_cruise_id, function(cruise_id){

  ## Cluster labels
  one_mapmat_long = mapmat_long %>% filter(cruise_id == !!(cruise_id)) %>% arrange(value) %>% na.omit()
  stopifnot((one_mapmat_long %>% pull(value)) == (1:10))
  cluster_labels = one_mapmat_long %>% pull(name)


  if(!is.null(objlist[[cruise_id]])){
    ## Alpha frequencies
    alpha_freq = objlist[[cruise_id]] %>% .$alpha_freq
    colnames(alpha_freq) = cluster_labels
    alpha_freq_long = alpha_freq %>% as.data.frame() %>% rownames_to_column("variable") %>% as_tibble() %>%
      pivot_longer(!variable, names_to = "population", values_to = "value") %>%
      add_column(type = "alpha")
  
    ## Beta frequencies
    beta_freq_long = lapply(1:3, function(idim){
      beta_array = objlist %>% .[[cruise_id]] %>% pluck("beta_freq") %>% .[,,idim] %>% t()
      colnames(beta_array) = cluster_labels
      one_long_beta = beta_array %>% as.data.frame() %>%
        rownames_to_column("variable")  %>% pivot_longer(!variable, names_to = "population") %>%
        add_column(type = paste0("beta_", cytogram_dimnames[idim]))
    }) %>% bind_rows() 
  
  
    ## Combine all frequencies
    freq_long = bind_rows(alpha_freq_long,
                          beta_freq_long)

  } else {
    freq_long = NULL
  }

  ## Beta coefficient values
  bestres = load_model(cruise_id)
  beta = bestres$beta
  names(beta) = cluster_labels
  beta = lapply(beta, function(onebeta){
    colnames(onebeta) = paste0("beta_", cytogram_dimnames); return(onebeta)
  })
  beta_long = lapply(cluster_labels, function(cluster_label){
    beta %>% .[[cluster_label]] %>% as.data.frame() %>%
      rownames_to_column("variable") %>% pivot_longer(!variable, names_to = "type") %>%
      add_column(population = cluster_label)
  }) %>% bind_rows()

  ## Alpha coefficients 
  alpha = bestres$alpha %>% t()
  colnames(alpha) = cluster_labels
  alpha_long = alpha %>% as.data.frame() %>%
    rownames_to_column("variable") %>%
    pivot_longer(!variable, names_to = "population") %>% 
    add_column(type = "alpha")

  ## Combine all coefficients
  coef_long = bind_rows(alpha_long,
                        beta_long)

  if(is.null(freq_long)) freq_long = coef_long[c(),]

  ## Combine everything
  combined = full_join(freq_long, coef_long, by=c("variable", "type", "population"), suffix = c("_stab", "_coef")) %>%
    select(variable, population, type, stab = value_stab, coef = value_coef)

  ## Add cruise_id and return
  combined = combined %>% add_column(cruise_id = cruise_id)

  return(combined)
}) %>% bind_rows() 
coef_long = coef_long %>% select(cruise_id, variable, population, type, coef, stab)
saveRDS(coef_long, file = file.path(outputdir, "coef_long_main.RDS"))
print(coef_long)
coef_long %>% head(20) %>% knitr::kable(caption = "Estimated coefficients in long format.")
```

**Predictions.** The columns are:

 + cruise
 + time
 + lon
 + lat
 + population
 + type ("mu_diam", "mu_chl", "mu_pe", "pi")
 + value (the value)
 


```{r pred_long}
pred_long = lapply(all_cruise_id, function(cruise_id){
  
  ## Cluster labels
  one_mapmat_long = mapmat_long %>% filter(cruise_id == !!(cruise_id)) %>% arrange(value) %>% na.omit()
  stopifnot((one_mapmat_long %>% pull(value)) == (1:10))
  cluster_labels = one_mapmat_long %>% pull(name)

  ## print(cruise_id)
  ## cluster_labels = NULL

  ## Times
  filename = paste0(cruise_id, "-datobj.RDS")
  ybin_obj = readRDS(file = file.path(outputdir_cytograms, cruise_id, "data", filename))

  ## Get time & lon & lat.
  times = ybin_obj$time %>% lubridate::as_datetime()
  lon = ybin_obj$lon
  lat = ybin_obj$lat
  
  ## Means
  bestres = load_model(cruise_id)
  mn_long = lapply(cytogram_dimnames, function(dimname){
    mn = bestres$mn %>% .[,which(cytogram_dimnames == dimname),]
    colnames(mn) = cluster_labels
    mn = mn %>% as_tibble() %>% add_column(time = times, lon = lon, lat = lat) %>%
      pivot_longer(-c(lon, lat, time), names_to = "population") %>% add_column(type=paste0("mu_", dimname))
    return(mn)
  }) %>% bind_rows() ##%>% pivot_wider(time, name)
  ## mn_wider = mn_long %>% pivot_wider(names_from = c("type"), values_from = "value")

  ## Probabilities
  prob = bestres$prob 
  colnames(prob) = cluster_labels
  prob_long = prob %>% as_tibble() %>% add_column(time = times, lon = lon, lat = lat) %>%
      pivot_longer(-c(lon, lat, time), names_to = "population") %>% add_column(type = "prob")


  ## Combine Prob and Means
  combined = bind_rows(prob_long, mn_long)##, sigma_long)

  ## Add cruise id and reaturn
  combined = combined %>% add_column(cruise_id = cruise_id)
  return(combined)
}) %>% bind_rows() 

pred_long = pred_long %>%  select(cruise_id, time, lon, lat, population, type, value)
saveRDS(pred_long, file = file.path(outputdir, "pred_long_main.RDS"))
pred_long %>% head(20) %>% knitr::kable(caption = "Predicted mean and abundance in long format.")
```


**Covariance matrix.** The columns are:

 + cruise
 + population
 + name: `"sigma_1_1"` means the (1,1)'th entry of the covariance matrix.
 + value: the value of the covariance matrix entry

```{r sigma_long}
sigma_long = sapply(all_cruise_id, function(cruise_id){

  ## Cluster labels
  one_mapmat_long = mapmat_long %>% filter(cruise_id == !!(cruise_id)) %>% arrange(value) %>% na.omit()
  stopifnot((one_mapmat_long %>% pull(value)) == (1:10))
  cluster_labels = one_mapmat_long %>% pull(name)

  ## Gather sigma
  sigmanames = paste0("sigma_", expand.grid(1:3, each=1:3) %>% apply(1, paste0, collapse="_"))
  sigma_wide = bestres$sigma %>% apply(1,c)
  colnames(sigma_wide) = cluster_labels 
  sigma_long = sigma_wide %>% as_tibble() %>% add_column(name = sigmanames) %>% pivot_longer(-name, names_to = "population")

  ## Also, here's a helper function to go from this to a 3 x 3 matrix
  if(FALSE){
    one_sigma_long = sigma_long %>% dplyr::filter(population == "pro1") ## this is an example
    make_covariance <- function(one_sigma_long){
      stopifnot(nrow(one_sigma_long) == 9)
      convertmat = one_sigma_long %>% pull(name) %>% sapply(., str_split, "_") %>% purrr::map(.%>%.[2:3]%>% as.numeric() ) %>% bind_rows() %>% t()
      sigmamat = matrix(NA, ncol = 3, nrow = 3)
      for(ii in 1:9){ 
        irow = convertmat[ii,1, drop=TRUE]
        icol = convertmat[ii,2, drop=TRUE]
        sigmamat[irow, icol] = one_sigma_long[ii, "value", drop=TRUE]
      }
      assertthat::assert_that(all(!is.na(sigmamat)))
      return(sigmamat)
    }
    make_covariance(one_sigma_long)
  }
  return(sigma_long)
}, simplify = FALSE, USE.NAMES = TRUE) %>% bind_rows(.id = "cruise_id") %>% arrange(cruise_id, population)

saveRDS(sigma_long, file = file.path(outputdir, "sigma_long_main.RDS"))
pred_long %>% head(20) %>% knitr::kable(caption = "Estimated covariance matrix in long format.")
```


**Covariates.** The columns are:

  + cruise
  + time
  + covariate
  + value

```{r X_long}
X_long = lapply(all_cruise_id, function(cruise_id){

  ## Get X
  bestres = load_model(cruise_id)
  X = bestres$X %>% as_tibble()

  ## Load times
  filename = paste0(cruise_id, "-datobj.RDS")
  ybin_obj = readRDS(file = file.path(outputdir_cytograms, cruise_id, "data", filename))
  ylist = ybin_obj$ylist
  times = names(ylist)

  ## Add time to X
  X = X %>% add_column(time = lubridate::as_datetime(times))


  ## Pivot longer
  X = X %>% pivot_longer(!time, names_to = "variable")

  ## Add cruise id
  X = X %>% add_column(cruise_id = cruise_id)

  return(X)
}) %>% bind_rows()

X_long = X_long %>% select(cruise_id, time, variable, value)
saveRDS(X_long, file = file.path(outputdir, "X_long_main.RDS"))
X_long %>% head(20) %>% knitr::kable(caption = "Environmental covariates in long format.")
```

**Data** A long matrix of binned data with the columns.

+ date/time
+ cruise_id
+ diam & chl & pe & biomass (for each bin that has nonzero data)
+ counts: biomass in that bin

```{r dat_long, eval = FALSE}
dat_list = sapply(all_cruise_id, function(cruise_id){
  ## Load cytogram data.
  filename = paste0(cruise_id, "-datobj.RDS")
  ybin_obj = readRDS(file = file.path(outputdir_cytograms, cruise_id, "data", filename))
  ylist = ybin_obj$ylist
  countslist = ybin_obj$biomass_list
  ylong = Map(function(y, c){data.frame(y, counts = c) %>% as_tibble()}, ylist, countslist) %>% bind_rows(.id = "time")
  return(ylong)
}, simplify = FALSE, USE.NAMES = TRUE)
dat_long = dat_list %>% bind_rows(.id = "cruise_id")
dat_long = dat_long %>% mutate(time = lubridate::as_datetime(time))
dat_long %>% object.size() %>% format("Mb")
saveRDS(dat_long, file = file.path(outputdir, "dat_long_main.RDS"))
dat_long %>% head(20) %>% knitr::kable(caption = "Binned cytometry data in long format.")
```

```{r knit-exit}
knitr::knit_exit()
```


# Next up

Finalize the pilot:

* Scale-free measure of each (covariate, cluster, cruise) to be compared across
  cruises. One idea is an ANOVA-type ratio of variance e.g. $var(X^T\beta_{salinity, k}) / var(\mu_k)$.
* Summarize (the absolute value of) coefficients across cruises.  
  + One plot per cluster; X axis is covariate value, and y axis is cruise.
* **done** Proper usage of random seeds (for experiments to be reproducible).
* **done** Calculate stability: 100x repeats of experiments.
* **done** Common-scaled covariate values (by fixing error in alpha M step).
* **done** Match and label a few clusters.
* **done** Match the clusters among the cruises, as much as possible, before producing
  these results. Label these clusters.
* **done** From the aggregated results, make plots with (y axis = which covariate) vs. (x
  axis = value).

Then, once the pilot is at a satisfactory state, do rest of experiments:

* Clean the cytograms from the remaining cruises.
* Run the jobs on the servers.
* Produce analyses.


Next things to do

* **DONE** Match the clusters among the cruises, as much as possible, before producing
  these results. Label these clusters.
* From the aggregated results, make plots with (y axis = which covariate) vs. (x
  axis = value).
* The estimated coefficients across cruises, in a single plot.

# Graveyard

## Notes from older experiments with non-common-scaled covariates 

Some notes of mine go here.

**Common observations**

* Persistent: two populations near prochloro's chl and diam, but one usually has
  no chl, and the pro has some chl.
* Two picoeuks that are separate.
* Synecho is nonexistent everywhere.

**KM1513**

* **Picoeuk(2&4)** Cluster 2 and 4 are separate picoeuk populations. Similar
  diameter, but 4 has lots of *pe*, while *2* have very little *pe*; both have
  *diam* and *pe* that oscillate with diel cycle, *together*.
* Continuing with 2 and 4; both seem to have a similar amount of Chl, but 4's
  *chl* oscillates much more than 2's *chl*.
* **Prochloro(6)** *10* and *6* seem to be separate populations; 6 is probably
  prochloro with lots of diel movement of diameter, while 10 doesn't. 6 has some
  *chl* (more than 10) while 10 has very small *chl*. Let's call 10 the "other".


**KOK1515**

* **Prochloro(10)** diel movement, and separate from 4.
* **Picoeuk(9&3)** 9 and 3 are the same as 2 and 4 from KM1513! Exciting -- look
  more carefully!

**KM1518**

* **Picoeuk(7&2)** same as 2&4 from KM1513.
* **Prochloro(10)** and a separate "other" 9.

**KM1512**

* **Picoeuk(3&1 and maybe 5)** It seems like 3&1 are 2&4 from KM1513 but 1
  captures a wider range in chl than in other cruises (there is less data here
  than in other cruises). We might try using even more clusters to narrow this down.
* **Prochloro(7&9)** 7&9 is prochloro, 10 is "other" (with no chl).

**KM1508**

* **Picoeuk(1&3 and maybe 4)** are same as 2&4. Cluster 2 has Strong diel
  moevement with diam and some pe as well. 3 has much smaller chl than in other
  cruises.
* **Prochloro(8&9)** CLEAR separation!
