---
title: Where does each species live?
date: "Compiled at `r format(Sys.time(), '%Y-%m-%d %H:%M:%S', tz = 'UTC')` UTC"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r global_options, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(fig.width=14, fig.height=8, echo=TRUE, eval=TRUE, cache=TRUE,
                      warning=FALSE, message=FALSE,
                      cache.lazy = FALSE)

## Load packages
library(tidyverse)
library(knitr)
library(here)
library(ggplot2)
library(gridExtra)
library(flowmix)
source("03-helpers.R")


## Setup some plotting details
coul <- colorRampPalette(RColorBrewer::brewer.pal(8, "RdYlBu"))(25)
## source("03-helpers.R")

## Some extra stuff, from: https://github.com/haozhu233/kableExtra/issues/265 
options(kableExtra.auto_format = F)
```

 
```{r directories}
## Setup locations
base = "03-where-plankton"
here::i_am("03-where-plankton.Rmd")
knitr::opts_chunk$set(fig.path = here::here("data", base, 'figures/'))
outputdir = here::here("data", base)
datadir = here::here("data", base)
tables_outputdir = here::here("data", "02-estimate")
cytograms_outputdir = here::here("data", "01-cytograms")
if(!dir.exists(outputdir)) dir.create(outputdir)
```

```{r data-setup}
pred_long_orig = readRDS(file = file.path(tables_outputdir, "pred_long_main.RDS"))

## Load the lons/lats
all_cruise_id = c("KM1508", "KM1512", "KM1513", "KOK1515", "KM1518", "KM1601",
                  "KM1602", "KOK1604", "KOK1606", "KOK1607", "KOK1608",
                  "KOK1609") 
spacetimes = lapply(all_cruise_id, function(cruise_id){
  ybin_obj = readRDS(file.path(cytograms_outputdir, cruise_id, "data", paste0(cruise_id, "-datobj.RDS")))
  tibble(cruise_id = cruise_id,
         time = ybin_obj$time,
         lon = ybin_obj$lon,
         lat = ybin_obj$lat)
}) %>% bind_rows()

## Add lon/lat to pred_long
pred_long = full_join(pred_long_orig, spacetimes)

## Only look at phytoplankton
phytopops = pred_long %>% pull(population) %>% unique() %>% grep(c("pro|syn|pico|croco"), ., value = TRUE)
pred_long = pred_long %>% dplyr::filter(population %in% phytopops)


## Collapse (1) pro1 and pro1_other into pro1.
## Collapse (2) croco and croco_other into croco.
## Collapse (3) pro1 and pico1-3 into pico.
pred_long = pred_long %>% combine_pop("pro1", "pro1")
pred_long = pred_long %>% combine_pop("croco", "croco")
pred_long_pico = pred_long %>% combine_pop("pico", "pico") %>% dplyr::filter(population == "pico") 
pred_long = rbind(pred_long, pred_long_pico)

## Reorder the populations
populations = pred_long$population %>% unique() %>% sort()
pred_long = pred_long %>% mutate(population = factor(population, levels = (populations)))
```



# Analysis goal

Parse the estimated estimated relative abundance $\pi$ from multiple cruises to
see where each species lives in the ocean.

# Prochlorococcus abundance

Let's focus on Prochlorococcus, for now. We'll be plotting **spatial averages**
of the relative abundance.

```{r plot-prochloro-overall, fig.width = 5, fig.height = 7}
p = pred_long %>%
  dplyr::filter(population == "pro1" & type == "prob") %>%
  average_by_space(cutwidth = 1)%>%
  plot_pipe(jitter_fac = 0, col_limits = c(0,.6)) +
  ggtitle("Prochloro abundance")
p + scale_size_continuous(range = rel(3), guide = "none") 
```

Spatial averages **by cruise** are shown here:

```{r plot-prochloro-by-cruise, fig.width = 15, fig.height = 15}
p = pred_long %>%
  dplyr::filter(population == "pro1" & type == "prob") %>%
  ## Obtain spatial averages /by cruise/
  split(.$cruise_id) %>%
  purrr::map_dfr(average_by_space, cutwidth = .5, .id = "cruise_id") %>%
  plot_pipe(jitter_fac = 0) +
  ggtitle("Prochloro abundance")
p + scale_size(range = rel(1), guide = "none") + facet_wrap(~cruise_id, nrow = 3)
```

Spatial averages **by cruise type** are shown here:

```{r plot-prochloro-by-cruise-type, fig.width = 12, fig.height = 6}
p = pred_long %>%
  left_join(tab) %>% 
  dplyr::filter(population == "pro1" & type == "prob") %>%
  ## Obtain spatial averages /by cruise/
  split(.$cruise_type) %>%
  purrr::map_dfr(average_by_space, cutwidth = .01, .id = "cruise_type") %>%
  ## plot_pipe(jitter_fac = 0)
  plot_pipe(jitter_fac = 0, col_limits = c(0,.6)) +
  ggtitle("Prochloro abundance")
p + scale_size(range = rel(1), guide = "none") + facet_wrap(~cruise_type, nrow = 1)
```


Now, we focus on latitude only.

* The points are colored by cruise.

* Each panel is a cruise types e.g. HOT, gradients.

```{r plot-avg-prochloro-by-latitude, fig.width = 12, fig.height = 6}
## Make an overall latitudinal average prob prediction.
latcut = pred_long %>%
  dplyr::filter(population == "pro1" & type == "prob") %>%
  mutate(latcut = ggplot2::cut_width(lat, width = .5)) %>%
  group_by(latcut) %>%
  summarize(value = mean(value),
            lat = mean(lat))
p = latcut %>% ggplot() +
  geom_line(aes(x=lat, y = value), size = rel(1), dat = latcut) +
  geom_point(aes(x=lat, y = value), size = rel(1), dat = latcut) +
  ggtitle("Average abundance by latitude") + ylab("Abundance") + xlab("Latitude")
p
```

What are the individual predictions?

```{r plot-prochloro-by-latitude, fig.width = 12, fig.height = 6}
## Overlay all points
ggplot() +
  geom_point(aes(x = lat, y = value), col = 'yellow',
             data = pred_long %>% filter(population == "pro1", type == "prob")) +
  geom_line(aes(x=lat, y = value), size = rel(1), dat = latcut, data = latcut) +
  ggtitle("Abundance by latitude") + ylab("Abundance") + xlab("Latitude")

## Overlay points and color by cruise ID
ggplot() +
  geom_point(aes(x = lat, y = value, col = cruise_id), alpha=1, size = rel(.5), 
             data = pred_long %>% filter(population == "pro1", type == "prob")) +
  guides(col = guide_legend(title = "Cruise ID")) +
  geom_line(aes(x=lat, y = value), size = rel(1), dat = latcut, data = latcut) +
  ggtitle("Abundance by latitude") + ylab("Abundance") + xlab("Latitude")

## Overlay points and color by cruise type
ggplot() +
  geom_point(aes(x = lat, y = value, col = cruise_type), alpha=.2,
             data = pred_long %>% left_join(tab) %>% filter(population == "pro1", type == "prob")) +
  guides(col = guide_legend(title = "Cruise type")) +
  geom_line(aes(x=lat, y = value), size = rel(1), dat = latcut, data = latcut) +
  ggtitle("Abundance by latitude") + ylab("Abundance") + xlab("Latitude")
```

Now, make one panel for each cruise type.

```{r plot-prochloro-cruisetype, fig.width = 20, fig.height = 10, eval = TRUE}
## Overlay points and color by cruise type
ggplot() +
  geom_point(aes(x = lat, y = value, col = cruise_type), alpha=.2,
             data = pred_long %>% left_join(tab) %>% filter(population == "pro1", type == "prob")) +
  guides(col = guide_legend(title = "Cruise type")) +
  facet_wrap(~cruise_type) +
  geom_line(aes(x=lat, y = value), size = rel(1), dat = latcut, data = latcut) +
  ggtitle("Abundance by latitude") + ylab("Abundance") + xlab("Latitude")
```

```{r plot-prochloro-cruisetype-zoomin, fig.width = 20, fig.height = 10, eval = TRUE}
cts = tab$cruise_type %>% unique() %>% sort()
glist = lapply(cts, function(ct){
## ct = "hot"
## ct = "hot"
dt = pred_long %>% 
  left_join(tab) %>% 
  filter(population == "pro1", type == "prob", cruise_type == ct)
rng = dt %>% pull(lat) %>% range()
  ylim = c(0, 0.7)
p = ggplot(dt) +
  geom_point(aes(x = lat, y = value, group = cruise_id, col = cruise_id)) +
  facet_wrap(~cruise_type) +
  ylim(ylim) +
  xlab("Latitude") + ylab("Abundance") +
  guides(col = guide_legend(title = "Cruise ID")) +
  ## Add the overall latitudinal prob average
  geom_line(aes(x=lat, y = value), size = rel(1), dat = latcut %>% filter(lat > rng[1] - .5 & lat < rng[2]+.5))
  return(p)
})
do.call("grid.arrange", glist)
```

If we just focus on the gradients(-ish) cruises:

```{r plot-prochloro-gradients, fig.width = 15, fig.height = 10}
p =
  pred_long %>%
  left_join(tab) %>%
  filter(population == "pro1", type == "prob", cruise_type == "gradients") %>%
  ggplot() +
  geom_point(aes(x = lat, y = value, group = cruise_id, col = cruise_id)) +
  xlab("Latitude") + ylab("Abundance") +
  guides(col = guide_legend(title = "Cruise ID")) +
  ## Add the overall latitudinal prob average 
  geom_line(aes(x=lat, y = value), size = rel(1), dat = latcut ) +
  facet_wrap(~cruise_id, nrow = 1) 
```

We see that, where the trajectories overlap, the three cruises agree quite
well. The trajectory for KM1713 differs substantially in the first part of the
cruise (which puts it in different nutrient conditions), which might be why the
KM1713's Synechococcus abundance is lower in latitudes below 30:

```{r plot-KM1713, fig.width = 7, fig.height = 5}
p =
  pred_long %>%
  filter(population == "pro1", type == "prob", cruise_id == "KM1713") %>%
  ggplot() +
  geom_point(aes(x = lat, y = value, group = cruise_id, col = cruise_id)) +
  xlab("Latitude") + ylab("Abundance") +
  guides(col = guide_legend(title = "Cruise ID")) +
  ## Add the overall latitudinal prob average 
  geom_line(aes(x=lat, y = value), size = rel(1), dat = latcut)
p

```

```{r plot-trajectory-gradients, fig.width = 5, fig.height = 8}
p = pred_long %>%
  left_join(tab) %>%
  filter(population == "pro1", type == "prob", cruise_type == "gradients") %>%
  ggplot() +
  geom_point(aes(x = lon, y = lat, group = cruise_id, col = cruise_id)) +
  xlab("Latitude") + ylab("Abundance") +
  guides(col = guide_legend(title = "Cruise ID")) +
  coord_map("azequalarea", orientation = c(-36.92, 200, 0)) +
  annotation_map(map_data("world"), fill = "antiquewhite", colour = "darkgrey") 
p
```

# Synecho & Croco abundance

We repeat this exercise for Synechococcus (and Croco) now.

```{r fig.height = 8, fig.width = 6}
pop = "syn"
p1 = pred_long %>%
  left_join(tab) %>% 
  dplyr::filter(population == pop & type == "prob") %>%
  average_by_space(cutwidth = .5)%>%
  plot_pipe(jitter_fac = 0) +
  ggtitle("Synechococcus abundance")  +
  scale_size(range = rel(3), guide = "none")

pop = "croco"
p2 = pred_long %>%
  left_join(tab) %>% 
  dplyr::filter(population == pop & type == "prob") %>%
  average_by_space(cutwidth = 1)%>%
  plot_pipe(jitter_fac = 0) +
  ggtitle("Crocosphaera abundance")  +
  scale_size_continuous(range = rel(3), guide = "none")

do.call("grid.arrange", list(p1, p2, nrow=1))
```

Synechococcus is 

(1) less abundant near the shore and up north, and
(2) more abundant in the middle.

Croco is more abundant in southern waters, and less abundant in northern waters.

Let's now focus on Synechococcus, and facet by cruise.

```{r, fig.width = 12, fig.height = 5}
pop = "syn"
p = pred_long %>%
  left_join(tab) %>% 
  dplyr::filter(population == pop & type == "prob") %>%
  ## Obtain spatial averages /by cruise/
  split(.$cruise_id) %>%
  purrr::map_dfr(average_by_space, cutwidth = .1, .id = "cruise_id") %>%
  plot_pipe(jitter_fac = 0, col_limits = c(0,.6)) +
  ggtitle("Synecho abundance")
p + scale_size_continuous(range = rel(1), guide = "none") + facet_wrap(~cruise_id, nrow = 1)
```
We'll focus on the two gradients cruises now.

It's important (and comforting) that Synecho relative abundance is *consistent*
on the two segments of the trip -- the way up, and the way down.

```{r synecho-gradients, fig.width = 8, fig.height=10}
pred_long %>%
  dplyr::filter(population == "syn" & type == "prob", cruise_id == "MGL1704") %>%
  mutate(segment = ifelse(time <= "2017-06-03 16:00:00", "up", "down")) %>%
  mutate(segment = factor(segment, levels=c("up", "down"))) %>% 
  plot_pipe(jitter_fac = 1) +
  facet_wrap(~segment, nrow = 1) +
  xlim(c(-155, -160)) +
  ggtitle("Synecho abundance in MGL1704")

pred_long %>%
  dplyr::filter(population == "syn" & type == "prob", cruise_id == "KOK1606") %>%
  mutate(segment = ifelse(time <= "2016-04-26 22:00:00", "up", "down")) %>%
  mutate(segment = factor(segment, levels=c("up", "down"))) %>% 
  plot_pipe(jitter_fac = 1) +
  facet_wrap(~segment, nrow = 1) +
  xlim(c(-155, -160)) +
  ggtitle("Synecho abundance in KOK1606")
```

```{r synecho-gradients-updown, fig.width = 8, fig.height = 5}
pred_long %>%
  dplyr::filter(population == "syn" & type == "prob", cruise_id == "MGL1704") %>%
  mutate(segment = ifelse(time <= "2017-06-03 16:00:00", "up", "down")) %>%
  mutate(segment = factor(segment, levels = c("up", "down"))) %>%
  ggplot() +
  geom_point(aes(x=lat, y=value, col = segment)) +
  ggtitle("Synecho abundance in MGL1704") + ylab("Abundance") + xlab("Latitude")

pred_long %>%
  dplyr::filter(population == "syn" & type == "prob", cruise_id == "KOK1606") %>%
  mutate(segment = ifelse(time <= "2016-04-26 22:00:00", "up", "down")) %>%
  mutate(segment = factor(segment, levels = c("up", "down"))) %>%
  ggplot() +
  geom_point(aes(x=lat, y=value, col = segment)) +
  ## geom_line(aes(x=lat, y=value, col = segment)) +
  ggtitle("Synecho abundance in KOK1606") + ylab("Abundance") + xlab("Latitude")
```

Next, we visualize the latitude-wise trend in Synechococcus abundance by (1)
cruise ID, and (2) cruise type.

```{r synecho-gradients-by-latitude, fig.width = 10, fig.height = 6}
latcut = pred_long %>%
  dplyr::filter(population == "syn" & type == "prob") %>%
  mutate(latcut = ggplot2::cut_width(lat, width = .5)) %>%
  group_by(latcut) %>%
  summarize(value = mean(value),
            lat = mean(lat))

pred_long %>%
  left_join(tab) %>% 
  dplyr::filter(population == "syn" & type == "prob") %>%
  ggplot() +
  geom_point(aes(x = lat, y = value, col = cruise_id)) +
  ggtitle("Synecho abundance, by cruise") + ylab("Abundance") + xlab("Latitude") +
  geom_line(aes(x = lat, y =value), dat = latcut, size = rel(2))

pred_long %>%
  left_join(tab) %>% 
  dplyr::filter(population == "syn" & type == "prob") %>%
  ggplot() +
  geom_point(aes(x = lat, y = value, col = cruise_type)) +
  ggtitle("Synecho abundance, by cruise type") + ylab("Abundance") + xlab("Latitude") +
  geom_line(aes(x = lat, y =value), dat = latcut, size = rel(2))
```

The two gradients cruises have similar *peak heights* (this is good) but the
latitude where the peaks occur differ quite a bit.  (KM1713 also went far up
north and down, but didn't have any Synechococcus detected; "bg" seems to be
where synecho usually is.)


```{r, fig.width = 7, fig.height = 5}
pred_long %>%
  left_join(tab) %>% 
  dplyr::filter(population == "syn" & type == "prob"& cruise_type == "gradients") %>%
  ggplot() +
  geom_point(aes(x = lat, y = value, col = cruise_id)) +
  ggtitle("Synecho abundance") + ylab("Abundance") + xlab("Latitude")
```

Why is this? The Transition zone actually migrates (fluctuates) seasonally, so
it's natural to find *seasonal variation* in the latitude at which the
Synechoccus thrives the most! (While there is also a temporal difference --
KOK1606 ran in late April and MGL ran in early June, this is not an important
distinction -- I'm told this is unimportant, and the main factor is a difference
in environmental conditions!)

Next up:

* Among HOT cruises, what does the data (covariates) say about the why they
  differ?

* Focus on environmental conditions and coefficients.

* Why are HOT cruises so variable?

* Incorporate gradients3 into the analysis.


```{r stop-eval, eval = TRUE}
knitr::knit_exit()   
```


<!-- ```{r colors} -->
<!-- ## colormap = setNames(RColorBrewer::brewer.pal(, "Set1") %>% adjustcolor(alpha = 0.5), nms) -->
<!-- colormap = c()  -->
<!-- colormap["light"] = "darkgoldenrod1" -->
<!-- colormap["phys"] = "firebrick" -->
<!-- colormap["bio"] = "springgreen4" -->
<!-- colormap["chem"] = "royalblue4" -->
<!-- colormap_small = colormap -->
<!-- colormap["all"] = "black" -->
<!-- colormap_ratio = colormap_small -->
<!-- names(colormap_ratio) = paste0(names(colormap_small), "_ratio") -->
<!-- ``` -->
